---
title: "R Notebook"
output: html_notebook
---

```{r}
# Libraries

# install.packages(igraph)
# install.packages('statnet')

library(tidyverse)
library(igraph)    # For creating the network
library(statnet)   # For ERGM
```

```{r}
# Baseline dataset provided by the professor of which names appear in which document
# ROWS: document
# COLS: names
#       - 1 : name appeared in document at that row
#       - 0 : name did not appear in document at that row
df <- read.csv('baseline/baseline_frequency.csv')

# Removing id column, changing colnames to numbers (for indexing later on),
# converting to matrix
df <- df %>% select(-1) 
mat <- as.matrix(df)

# Loop for populating our adjacency matrix
adj.mat <- matrix(0, ncol(mat), ncol(mat))
for (doc in 1:nrow(mat)) {
  for (i in 1:ncol(mat)) {
    if (mat[doc, i] != 0) {
      for (j in i+1:length(mat[doc,])) {
        # R sucks and if i+1 == 109, the loop guard doesn't care and it will 
        # just count backwards down to 108.
        # So you need this extra if statement.
        if (j < 109) {
          if (mat[doc, j] != 0) {
            adj.mat[i, j] <- 1
            adj.mat[j, i] <- 1
          }
        }
      }
    }
  }
}
```

```{r}
n <- network(adj.mat, directed=FALSE)

# Set east vs. west attribute for each node
names <- read.csv('names.csv')
names.filtered <- names[!duplicated(names[,c('lastname')]),][1:108,]
names.filtered.sorted <- names.filtered[order(names.filtered$lastname),]

names.filtered.sorted

# # Checking that names are in same order in both vectors
n %v% 'vertex.names'<- str_to_title(sort(names.filtered.sorted['lastname']))
#network.vertex.names(g)

#df
#names
#names.filtered[,'lastname']

for (i in 1:108) {
  print(paste(i, paste(names.filtered.sorted[i,'lastname'], colnames(df)[i]))) }

#names.filtered[,'affiliation']

n%v%'vertex.names'<- colnames(df)
n%v%'affiliation'<- names.filtered.sorted[,'affiliation']
n%v%'DC' <- names.filtered.sorted[,'DC.Y.N.']
n%v%'media' <- names.filtered.sorted[,'media.Y.N.']
n%v%'private' <- names.filtered.sorted[,'private.Y.N.']

# network.vertex.names(n)
# plot(n)
```

## Model 1: Null model

```{r}
# NULL MODEL
# Estimate the simplest model, one with only a term for tie density (akin to an intercept term in a glm)
null.model = ergm(n ~ edges)
summary(null.model)
```

The conditional log-odds of two actors having a tie, keeping the rest of the network fixed, is -2.9332 * change in the number of ties.

## Model 2: Triad Model

Let’s add a term often thought to be a measure of “clustering”: the number of completed triangles in the network.

```{r}
## TAKES FOREVER TO RUN, HAVE NO CLUE WHAT THE STATS FOR THIS ONE ARE
#model.2 <- ergm(n ~ edges + gwesp)
#summary(model.2)
?ergm.terms
```

## Model 3: Affiliation (East vs. West)

```{r}
# affiliation.model = ergm(n ~ edges + nodematch('affiliation', diff=FALSE, keep=NULL, levels=NULL))
affiliation.model = ergm(n ~ edges + nodefactor('affiliation', base=1, levels=-1) 
                         + nodematch('affiliation', diff=FALSE, keep=NULL, levels=NULL))

summary(affiliation.model)
```

The conditional log-odds of two actors having a tie, keeping the rest of the network fixed, is -3.8994 * change in the number of ties + 0.16382 if node in question is a westerner + 1.33859 if both nodes share the same affiliation.\

DOUBLE CHECK send/receive -- .32?

```{r}
# affiliation.model = ergm(n ~ edges + nodematch('affiliation', diff=FALSE, keep=NULL, levels=NULL))
affiliation.model = ergm(n ~ edges + nodefactor('DC', base=1, levels=-1))

summary(affiliation.model)
```

```{r}
# affiliation.model = ergm(n ~ edges + nodematch('affiliation', diff=FALSE, keep=NULL, levels=NULL))
affiliation.model = ergm(n ~ edges + nodefactor('DC', base=1, levels=-1))

summary(affiliation.model)
```

```{r}
# affiliation.model = ergm(n ~ edges + nodematch('affiliation', diff=FALSE, keep=NULL, levels=NULL))
affiliation.model = ergm(n ~ edges + nodefactor('media', base=1, levels=-1))

summary(affiliation.model)
```


```{r}
# affiliation.model = ergm(n ~ edges + nodematch('affiliation', diff=FALSE, keep=NULL, levels=NULL))
affiliation.model = ergm(n ~ edges + nodefactor('private', base=1, levels=-1))

summary(affiliation.model)
```
