---
title: "R Notebook"
output: html_notebook
---

```{r}
# Libraries

# install.packages(igraph)
# install.packages('statnet')

library(tidyverse)
library(igraph)    # For creating the network
library(statnet)   # For ERGM
```

```{r}
# Baseline dataset provided by the professor of which names appear in which document
# ROWS: document
# COLS: names
#       - 1 : name appeared in document at that row
#       - 0 : name did not appear in document at that row
df <- read.csv('baseline/baseline_frequency.csv')

# Removing id column, changing colnames to numbers (for indexing later on),
# converting to matrix
df <- df %>% select(-1) 
mat <- as.matrix(df)

# Loop for populating our adjacency matrix
adj.mat <- matrix(0, ncol(mat), ncol(mat))
for (doc in 1:nrow(mat)) {
  for (i in 1:ncol(mat)) {
    if (mat[doc, i] != 0) {
      for (j in i+1:length(mat[doc,])) {
        # R sucks and if i+1 == 109, the loop guard doesn't care and it will 
        # just count backwards down to 108.
        # So you need this extra if statement.
        if (j < 109) {
          if (mat[doc, j] != 0) {
            adj.mat[i, j] <- 1
            adj.mat[j, i] <- 1
          }
        }
      }
    }
  }
}
```

```{r}
# Set east vs. west attribute for each node
names <- read.csv('names_copy.csv')
```

```{r}
n <- network(adj.mat, directed=FALSE)

names.filtered <- names[!duplicated(names[,c('lastname')]),][1:108,]

# # Checking that names are in same order in both vectors
# tolower(names.filtered[,'lastname'])
# network.vertex.names(n)
# tolower(network.vertex.names(n)) == tolower(names.filtered[,'lastname'])

names.filtered[,'affiliation']

affil.binary <- gsub('east', '1', names.filtered[,'affiliation'])
affil.binary <- gsub('west', '0', affil.binary)
affil.binary <- as.integer(affil.binary)

affil.binary

n%v%'vertex.names'<- colnames(df)
n%v%'affiliation'<- names.filtered[,'affiliation']
# network.vertex.names(n)
# plot(n)
```

## Model 1: Null model

```{r}
# NULL MODEL
# Estimate the simplest model, one with only a term for tie density (akin to an intercept term in a glm)
null.model = ergm(n ~ edges)
summary(null.model)
```

The conditional log-odds of two actors having a tie, keeping the rest of the network fixed, is -2.9332 * change in the number of ties.

## Model 2: Triad Model

Let’s add a term often thought to be a measure of “clustering”: the number of completed triangles in the network.

```{r}
## TAKES FOREVER TO RUN, HAVE NO CLUE WHAT THE STATS FOR THIS ONE ARE
model.1 <- ergm(n ~ edges + triangle)
summary(model.1)
```

## Model 3: Affiliation (East vs. West)

```{r}
affiliation.model = ergm(n ~ edges + nodematch('affiliation'))
summary(affiliation.model)
```

The conditional log-odds of two actors having a tie, keeping the rest of the network fixed, is -3.8994 * change in the number of ties + 1.4271 * {1 if node 1 is an easterner} + 1.4271 * {1 if node 2 is an easterner}.


